Demand Forecasting of Individual Probability Density Functions with Machine Learning

### 2020-12-21 LaTex/paper.tex
several changes after review

% quantile regression
Generally, quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Focusing on time series, Hyndman \cite{hyndman1995highest} proposes to use either a simulation or bootstrapping framework to estimate the highest density forecast regions of non-normal time series. Tay \cite{tay2000density} summarizes the use of density forecasts in applications in finance and economics. Using RNNs, quantile regression for multiple forecasting horizons \cite{wen2017multi} has been used to predict the demand at the internet retailer Amazon. Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step.

@@

A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted PDF $f(x)$ is correctly calibrated. This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed CDF values for five different cases. If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution. If the mean or the variance are not estimated correctly, the resulting distribution will show a distinct deviation from this uniform behavior.

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{figs/cdf_truth_uniform}
\includegraphics[scale=0.25]{figs/cdf_truth_broad}
\includegraphics[scale=0.25]{figs/cdf_truth_narrow}
\includegraphics[scale=0.25]{figs/cdf_truth_over}
\includegraphics[scale=0.25]{figs/cdf_truth_under}
\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
correct prediction (a: ``uniform''), variance overestimated (b: ``broad''), variance underestimated (c: ``narrow''), mean overestimated (d: ``over''), mean underestimated (e: ``under'')}
\end{center}
\end{figure}

@@

\noindent
As discussed in sec. \ref{sec:ts_confounding}, we do not use any lagged target information in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with a recursive smoothing factor of $0.15$) of the predictions and targets of each product-location combination over the corresponding past. To reflect a realistic replenishment scenario, we use the model to predict the expected demand two days into the future, meaning that we use a lag of two days for the calculation of  both the target EWMA and the prediction EWMA.

\noindent
We use the mean absolute deviation (MAD) as well as the mean squared error (MSE) as two common metrics for the evaluation of point estimates to give a rough estimate of the accuracy of the predicted mean. These metrics do not take the shape of the underlying PDF into account, but only compare the predicted mean to the observed number of sales. Using the independent test data, we obtain the following metrics: MAD: $1.65$ and MSE: $10.09$. The mean of the target, i.e., the observed sales, is $3.28$ for this period. \fig \ref{fig:mean_prediction} shows the time series of both predictions and sales summed over all 100 products and 10 stores during the test period.

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{figs/ts_item_16_store_2_res}
\caption{\label{fig:ts_res} Time series of mean predictions, before (machine learning prediction) and after (final prediction) residual correction, as well as sales for a specific product-location combination. In addition to the test period in 2016, the last two months of the training period are shown as well, in order to better demonstrate the workings of the residual correction.}
\end{center}
\end{figure}

\noindent
To give an indication of the effectiveness of the two-step approach, i.e., machine learning model without lagged target information and residual correction, we get the following metrics for a model without residual correction, but including the EWMA of the target (grouped by \texttt{store\_id} and \texttt{item\_id}, with a lag of two days, and a recursive smoothing factor of $0.25$) as a feature: MAD: $1.70$ and MSE: $11.04$. The value of the smoothing factor here differs from the one used in the residual correction, because this stricter damping improved the evaluation metrics. Including the day of the week as an additional group by dimension in the EWMA feature worsened the metrics even more.

>>>>

% quantile regression
Generally, quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated. Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model. Focusing on time series, Hyndman \cite{hyndman1995highest} proposes to use either a simulation or bootstrapping framework to estimate the highest density forecast regions of non-normal time series. Tay \cite{tay2000density} summarizes the use of density forecasts in applications in finance and economics. Using RNNs, quantile regression for multiple forecasting horizons \cite{wen2017multi} has been used to predict the demand at the internet retailer Amazon. Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step. Another possibility to model the underlying data distribution without assumption of a fixed distribution class is via conditional normalizing flows \cite{rezende2016variational,timeseriesflow}, which aim to transform a simple initial density into a more complex one by applying a sequence of invertible transformations.

@@

A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted PDF $f(x)$ is correctly calibrated, i.e., if both the choice of the model and the model parameters are estimated correctly. If the mean or the variance are not estimated correctly, the resulting distribution will show a distinct deviation from this uniform behavior. This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed CDF values for four different cases.

\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{figs/cdf_truth_broad}
\includegraphics[scale=0.25]{figs/cdf_truth_narrow}
\includegraphics[scale=0.25]{figs/cdf_truth_over}
\includegraphics[scale=0.25]{figs/cdf_truth_under}
\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution (dashed horizontal line in each of the plots): variance overestimated (a: ``broad''), variance underestimated (b: ``narrow''), mean overestimated (c: ``over''), mean underestimated (d: ``under'')}
\end{center}
\end{figure}

@@

\noindent
As discussed in sec. \ref{sec:ts_confounding}, we do not use any lagged target information in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with $\alpha=0.15$) of the predictions and targets of each product-location combination over the corresponding past. Empirically, a value of $\alpha=0.15$ gives good results here, though in a practical application, this parameter would have to be optimized using methods such as cross-validation according to a metric that reflects the objective of the business strategy. To reflect a realistic replenishment scenario, we use the model to predict the expected demand two days into the future, meaning that we use a lag of two days for the calculation of both the target EWMA and the prediction EWMA.

\noindent
We use the mean absolute deviation (MAD) as well as the mean squared error (MSE) as two common metrics for the evaluation of point estimates to give a rough estimate of the accuracy of the predicted mean. These metrics do not take the shape of the underlying PDF into account, but only compare the predicted mean to the observed number of sales. Using the independent test data, we obtain the metrics as stated in the top line of \tab \ref{tab:mad_mse}. The mean of the target, i.e., the observed sales, is $3.28$ for this period. \fig \ref{fig:mean_prediction} shows the time series of both predictions and sales summed over all 100 products and 10 stores during the test period.

\begin{table}[h!]
\begin{center}
\caption{Accuracy metrics for the predicted mean values for three different model setups: (a) ML model without lagged target information but subsequent residual correction, (b) ML model including lagged target information by means of two EMWA features without residual correction, and (c) ML model from (b) with residual correction.}
\label{tab:mad_mse}
\begin{tabular}{p{4cm}|c|c}
 & \textbf{MAD} & \textbf{MSE} \\
\hline
\textbf{no EWMA features,\newline residual correction} & 1.65 & 10.09 \\
\hline
\textbf{EWMA features,\newline no residual correction} & 1.69 & 10.84 \\
\hline
\textbf{EWMA features,\newline residual correction} & 1.68 & 10.41
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{figs/ts_item_16_store_2_res}
\caption{\label{fig:ts_res} Time series of mean predictions, before (machine learning prediction) and after (final prediction) residual correction, as well as sales for a specific product-location combination. In addition to the test period in 2016, the last two months of the training period are shown as well (the dashed vertical line marking the transition from training to test period), in order to better demonstrate the workings of the residual correction.}
\end{center}
\end{figure}

\noindent
To give an indication of the effectiveness of the two-step approach, i.e., a machine learning model without lagged target information and residual correction as outlined in sec. \ref{sec:ts_confounding}, we compare its results (for the same training and test setup) with a model following the common practice of including the lagged target information as a variety of EMWA variables used as features in \tab \ref{tab:mad_mse}. For this, we include two EWMA features in the machine learning model, one grouped by \texttt{store\_id} and \texttt{item\_id} (with a lag of two days and $\alpha=0.25$), and one grouped by \texttt{store\_id}, \texttt{item\_id}, and the day of the week (with a lag of one week and $\alpha=0.05$), where the values for $\alpha$ are again found empirically. In \tab \ref{tab:mad_mse}, we also show the results after applying the same residual correction as before for the machine learning model without lagged target information on top of this model including the two EWMA features. The crucial point in this is that when applying the same residual correction method to both machine learning models with and without including lagged target information as EWMA features, the one without yields better results. Following our reasoning in sec. \ref{sec:ts_confounding}, this is due to detrimental temporal confounding effects disturbing the learning of the causal dependencies between the exogenous features and the target, which in turn leads to a worse generalizability of the machine learning model.
